Conference Paper Title\\

### 2023-06-17 LARS_paper/paper.tex
Fix pontuation and minor issues

The robot platform designed for this research is presented in the Fig. \ref{fig:robot}. The development encompassed the prototyping and the implementation of the gait planner, the trajectory planner, the kinematic model, and the control subsystem.

\begin{figure}[!htb]
  \centering
  \includegraphics[width=0.25\textwidth]{robot.jpeg}
  \vfill
  % \vspace{-\baselineskip}
  \caption{The quadruped robot}
  \label{fig:robot}
\end{figure}

@@

\begin{figure}[!htb]
  \centering
  \includegraphics[width=0.45\textwidth]{trot_pattern.png}
  \vfill
  % \vspace{-\baselineskip}
  \caption{Gait pattern diagram}
  \label{fig:trot_pattern}
\end{figure}

@@

\begin{table}[!htb]
  \centering
  \caption{Results of the gait trajectory tracking experiment.}
  \scalebox{1.3}{
  \begin{tabular}{ccc}
    \hline
    \textbf{Test} & \textbf{1}      & \textbf{2}  \\
    \hline
    $\bar{time} (s)$         & 0,557  & 0,557 \\
    \hline
    $\sigma_{time} (s)$       & 0,008  & 0,008 \\
    \hline
    $\bar{z}_{max} (cm)$       & 4,760  & 4,755 \\      
    \hline
    $\sigma_{z_{max}} (cm)$    & 0,023  & 0,028 \\
    \hline
    $\bar{x}_{final} (cm)$     & 5,034  & 3,024 \\
    \hline
    $\sigma_{x_{final}} (cm)$  & 0,009  & 0,013 \\
    \hline
    $\bar{y}_{final} (cm)$     & 2,924  & 4,915 \\      
    \hline
    $\sigma_{y_{final}} (cm)$  & 0,013  & 0,004 \\      
    \hline   
  \end{tabular}
  }
  \vspace{-\baselineskip}
  \label{tab:trajetoria}
\end{table}

\begin{figure}[!htb]
  \centering
  \caption{Trajetórias realizadas pelas patas.}
  \includegraphics[width=0.48\textwidth]{grafico_trajetoria_z.png}
  \label{fig:grafico_trajetoria_xyz}
\end{figure}
\vspace{-\baselineskip}

It is noticeable that there was a delay of approximately 57 ms in the execution time and that the final height was off by approximately 0.25 cm in both tests. To confirm that the system's performance during Test 1 and 2 was similar, i.e. the difference in x and y distance did not affect the overall performance, two ANOVA \cite{cano2012six} tests were conducted comparing the execution time and gait height distributions of both tests. Before the ANOVA, however, the outliers of these distributions were removed and the Shapiro-Wilk \cite{leotti2005comparaccao} test was used to check the data normality. The result of the later was a p-value greater than 0.05 for all distributions, which indicates that they followed a normal curve. The ANOVA tests were, then, conducted and showed a p-value of 0.7212 for the execution time comparison and 0.5070 for the gait height comparison, which means that de difference in the x and y distance did not affect the trajectory tracking performance. In addition to that, the final average error for x and y was less than 0.1 cm and the standard deviation was small, meaning the gait distance was followed correctly and that this result was consistent throughout all samples.

@@

\begin{figure}[htbp]
  \centering
  \begin{subfigure}[htbp]{0.24\textwidth}
    \centering
    \includegraphics[width=1.0\textwidth]{terreno_plano.jpeg}
    \caption{Flat.}
    \label{fig:terreno_plano}
  \end{subfigure}
  \begin{subfigure}[htbp]{0.24\textwidth}
    \centering
    \includegraphics[width=1.0\textwidth]{terreno_irregular.jpeg}
    \caption{Uneven.}
    \label{fig:terreno_uneven}
  \end{subfigure}
  \vfill
  \caption{Terrains used for the experiment}
  \label{fig:terrenos}
\end{figure}

@@

\begin{table}[!htb]
  \centering
  \caption{Velocity and stability results}
  \scalebox{1.3}{
  \begin{tabular}{ccccc}
    \hline
    \textbf{Test}              & \textbf{1}   & \textbf{2}  & \textbf{3}   & \textbf{4}   \\ \hline
    Ter.                       & Flat         & Flat        & Uneven       & Uneven       \\ \hline
    \begin{tabular}[c]{@{}c@{}}S. C. \end{tabular} & No           & Yes         & No           & Yes          \\ \hline
    \begin{tabular}[c]{@{}c@{}}Vel. \\ $(cm/s)$ \end{tabular} & 2.15         & 3.68        & 2.03         & 2.39         \\ \hline
    \begin{tabular}[c]{@{}c@{}} $\sigma_{Vel}$  \\ $(cm/s)$ \end{tabular} & 0.040        & 0.075       & 0.016        & 0.047        \\ \hline
    \begin{tabular}[c]{@{}c@{}} $\Delta_{Roll}$ \end{tabular} & 12.59\degree & 8.81\degree & 13.37\degree & 11.13\degree \\\hline
    \begin{tabular}[c]{@{}c@{}} $\sigma_{Roll}$ \end{tabular} & 2.54\degree  & 2.36\degree & 0.78\degree  & 1.47\degree  \\ \hline
    \begin{tabular}[c]{@{}c@{}} $\Delta_{Pitch}$ \end{tabular} & 11.55\degree & 8.31\degree & 15.19\degree & 12.09\degree \\ \hline
    \begin{tabular}[c]{@{}c@{}} $\sigma_{Pitch}$ \end{tabular} & 1.89\degree  & 2.63\degree & 1.44\degree  & 1.66\degree  \\ \hline
  \end{tabular}
  }
  \label{tab:vel_stab}
\end{table}

>>>>

The robot platform designed for this research is presented in Fig. \ref{fig:robot}. The development encompassed the prototyping and the implementation of the gait planner, the trajectory planner, the kinematic model, and the control subsystem.

\begin{figure}[!htb]
  \centering
  \includegraphics[width=0.4\textwidth]{robot.jpeg}
  \vfill
  % \vspace{-\baselineskip}
  \caption{The quadruped robot.}
  \label{fig:robot}
\end{figure}

@@

\begin{figure}[!htb]
  \centering
  \includegraphics[width=0.45\textwidth]{trot_pattern.png}
  \vfill
  % \vspace{-\baselineskip}
  \caption{Gait pattern diagram.}
  \label{fig:trot_pattern}
\end{figure}

@@

\begin{table}[!htb]
  \centering
  \caption{Results of the gait trajectory tracking experiment.}
  \scalebox{1.3}{
  \begin{tabular}{ccc}
    \hline
    \textbf{Test} & \textbf{1}      & \textbf{2}  \\
    \hline
    $\bar{time} (s)$         & 0.557  & 0.557 \\
    \hline
    $\sigma_{time} (s)$       & 0.008  & 0.008 \\
    \hline
    $\bar{z}_{max} (cm)$       & 4.760  & 4.755 \\      
    \hline
    $\sigma_{z_{max}} (cm)$    & 0.023  & 0.028 \\
    \hline
    $\bar{x}_{final} (cm)$     & 5.034  & 3.024 \\
    \hline
    $\sigma_{x_{final}} (cm)$  & 0.009  & 0.013 \\
    \hline
    $\bar{y}_{final} (cm)$     & 2.924  & 4.915 \\      
    \hline
    $\sigma_{y_{final}} (cm)$  & 0.013  & 0.004 \\      
    \hline   
  \end{tabular}
  }
  \vspace{-\baselineskip}
  \label{tab:trajetoria}
\end{table}

\begin{figure}[!htb]
  \centering
  \includegraphics[width=0.48\textwidth]{grafico_trajetoria_z.png}
  \caption{Trajectory executed by the foot.}
  \label{fig:grafico_trajetoria_xyz}
\end{figure}
\vspace{-\baselineskip}

It is noticeable that there was a delay of approximately 57 ms in the execution time and that the final height was off by approximately 0.25 cm in both tests. To confirm that the system's performance during Test 1 and 2 was similar, i.e. the difference in x and y distance did not affect the overall performance, two ANOVA \cite{cano2012six} tests were conducted comparing the execution time and gait height distributions of both tests. Before the ANOVA, however, the outliers of these distributions were removed and the Shapiro-Wilk \cite{leotti2005comparaccao} test was used to check the data normality. The result of the latter was a p-value greater than 0.05 for all distributions, which indicates that they followed a normal curve. The ANOVA tests were, then, conducted and showed a p-value of 0.7212 for the execution time comparison and 0.5070 for the gait height comparison, which means that de difference in the x and y distance did not affect the trajectory tracking performance. In addition to that, the final average error for x and y was less than 0.1 cm and the standard deviation was small, meaning the gait distance was followed correctly and that this result was consistent throughout all samples.

@@

\begin{figure}[htbp]
  \centering
  \begin{subfigure}[htbp]{0.24\textwidth}
    \centering
    \includegraphics[width=1.0\textwidth]{terreno_plano.jpeg}
    \caption{Flat.}
    \label{fig:terreno_plano}
  \end{subfigure}
  \begin{subfigure}[htbp]{0.24\textwidth}
    \centering
    \includegraphics[width=1.0\textwidth]{terreno_irregular.jpeg}
    \caption{Uneven.}
    \label{fig:terreno_uneven}
  \end{subfigure}
  \vfill
  \caption{Terrains used for the experiment.}
  \label{fig:terrenos}
\end{figure}

@@

\begin{table}[!htb]
  \centering
  \caption{Velocity and stability results.}
  \scalebox{1.3}{
  \begin{tabular}{ccccc}
    \hline
    \textbf{Test}              & \textbf{1}   & \textbf{2}  & \textbf{3}   & \textbf{4}   \\ \hline
    Ter.                       & Flat         & Flat        & Uneven       & Uneven       \\ \hline
    \begin{tabular}[c]{@{}c@{}}S. C. \end{tabular} & No           & Yes         & No           & Yes          \\ \hline
    \begin{tabular}[c]{@{}c@{}}Vel. \\ $(cm/s)$ \end{tabular} & 2.15         & 3.68        & 2.03         & 2.39         \\ \hline
    \begin{tabular}[c]{@{}c@{}} $\sigma_{Vel}$  \\ $(cm/s)$ \end{tabular} & 0.040        & 0.075       & 0.016        & 0.047        \\ \hline
    \begin{tabular}[c]{@{}c@{}} $\Delta_{Roll}$ \end{tabular} & 12.59\degree & 8.81\degree & 13.37\degree & 11.13\degree \\\hline
    \begin{tabular}[c]{@{}c@{}} $\sigma_{Roll}$ \end{tabular} & 2.54\degree  & 2.36\degree & 0.78\degree  & 1.47\degree  \\ \hline
    \begin{tabular}[c]{@{}c@{}} $\Delta_{Pitch}$ \end{tabular} & 11.55\degree & 8.31\degree & 15.19\degree & 12.09\degree \\ \hline
    \begin{tabular}[c]{@{}c@{}} $\sigma_{Pitch}$ \end{tabular} & 1.89\degree  & 2.63\degree & 1.44\degree  & 1.66\degree  \\ \hline
  \end{tabular}
  }
  \label{tab:vel_stab}
\end{table}
